Quantizing the Bosonic String on a Loop Quantum Gravity Background

### 2024-04-15 lqg-strings.tex
sketched some stuff and started correcting the main file

In the same way we considered the quantum geometry inverse metric correction to the Nambu-Goto action, it is also natural to consider such correction to the Polyakov string. we start with the action(\ref{eqn:polyakov-gravity}) and proceed with the inverse area correction(\ref{eqn:inverse-metric-correction}), which yields our corrected Polyakov action
\begin{multline} \label{eqn:inverse-metric-Polyakov}
	S_{I A P} = - \frac{T}{2} \int \diff ^2 x \left( \sqrt{- g} - \frac{k \Delta}{2 \sqrt{- g}} \right) \times \\ \times g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu} .
\end{multline}
The equations of motion for the world-sheet metric are completely unchanged by this correction, so they are
\begin{equation} \label{eqn:vary-wrt-g}
	\frac{\delta S_{I A P}}{\delta g^{c d}} \overset{!}{=} 0
\end{equation}
from which we obtain the stress-energy tensor for the worldsheet:
\begin{multline} \label{eqn:inverse-area-em-tensor}
	T_{c d} := \partial _{c} X^{\mu} \partial _{d} X^{\nu} G_{\mu \nu} - \\ - \frac{1}{2} g_{c d} g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu} = 0 ,
\end{multline}
from which we can extract the metric $g_{c d}$ as
\begin{equation} \label{eqn:ws-metric}
	g_{c d} = 2 f \partial _{c} X^{\mu} \partial _{d} X^{\nu} G_{\mu \nu} ,
\end{equation}
where the function $f$ is given by
\begin{equation} \label{eqn:inverse-area-f}
	\frac{1}{f} = g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu} .
\end{equation}

As for variation w.r.t the embbeding fields $X^{\lambda}$, we have
\begin{equation} \label{eqn:vary-wrt-X}
	\frac{\delta S_{I A P}}{\delta X^{\lambda}} \overset{!}{=} 0
\end{equation}

\begin{multline} \label{eqn:X-eom-inverse-area}
	F^{-} \sqrt{- g} g^{a b} \partial _{a} \partial _{b} X^{\nu} G_{\lambda \nu} + \partial _{a} \left( F^{-} \sqrt{- g} g^{a b} \right) \partial _{b} X^{\nu} G_{\lambda \nu} = \\ = \frac{1}{2} F^{-} \sqrt{- g} g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} \partial _{\lambda} G_{\mu \nu} ,
\end{multline}
where $F^{-} = 1 - k \Delta / 2 (- g)$. Using conformal symmetry $g^{a b} = (\phi (x))^{-1} \eta ^{a b}$ and assuming flat background $G_{\mu \nu} = \eta _{\mu \nu}$, eq(\ref{eqn:X-eom-inverse-area}) simplify to
\begin{equation} \label{eqn:X-eom-simplified}
	\eta ^{a b} \partial _{a} \partial _{b} X^{\mu} + \eta ^{a b} \partial _{a} \ln (F^{-}) \partial _{b} X^{\mu} = 0 .
\end{equation}

@@

Equation(\ref{eqn:X-eom-simplified}) seems a bit odd, but we can polish it by choosing a plane-wave ansatz for the embedding fields $X^{\mu} = X^{\mu} _{0} \e ^{-i (E \tau - p \sigma)}$ and using it to fix a form for the conformal function $\phi (x)$ appearing in $F^{-}$. By using the ansatz in eq(\ref{eqn:X-eom-simplified}), it becomes
\begin{equation}
	(E^{2} - p^{2}) X^{\mu} + i (E \partial _{\tau} \ln (F^{-}) - p \partial _{\sigma} \ln (F^{-})) X^{\mu} = 0 ,
\end{equation}
from which we can use the freedom in the conformal function to fix the second term as $- m^2 X^{\mu}$, with $- m^2 = p^2 - E^2$ on-shell, and this in turn implies that the derivatives of $\ln (F^{-})$ are given by
\begin{equation} \label{eqn:derivatives-F-minus}
	\partial _{\tau} \ln (F^{-}) = i E , \ \partial _{\sigma} \ln (F^{-}) = i p ,
\end{equation}
thus we have
\begin{equation} \label{eqn:ln-F}
	\ln (F^{-}) = i (E \tau + p \sigma) + a , \ a \in \mathbb{C}
\end{equation}

\begin{equation} \label{eqn:F-minus-exp}
	F^{-} = A \e ^{i (E \tau + p \sigma + \varphi)} , \ A > 0
\end{equation}
\begin{equation} \label{eqn:phi}
	\phi (\tau , \sigma) = \sqrt{\frac{k \Delta}{2 \left( 1 - A \e ^{i (E \tau + p \sigma + \varphi)} \right)}} .
\end{equation}
At first glance it seems problematic that we have a complex conformal function $\phi (x)$, but the complex structure comes from our ansatz for the embedding fields, and when we choose either the real or imaginary part of the latter it should also pick the respective component of $\phi$. With this choice, we thus turned eq(\ref{eqn:X-eom-simplified}) into a Klein-Gordon equation
\begin{equation} \label{eqn:X-Klein-Gordon-eqn}
	\left( \partial _{\sigma} ^2 - \partial _{\tau} ^2 \right) X^{\mu} - \mu ^2 X^{\mu} = 0 .
\end{equation}

@@

At first glance, our recent analysis of the Polyakov string with inverse metric correction seems to not be related to what we found in the analysis of the Nambu-Goto string. However, as pointed out in \cite{Grimshaw2010Homogenization}, wave equations with a variable speed (the result of the Nambu-Goto analysis) can be turned into Klein-Gordon equations (what we just found). While they're work deals only with space-varying wave speed, it is straight forward to generalise the procedure to also include time dependence, but ironically enough the calculations lead to the conclusion that the wave speed $v = 1 / F^{+}$ is actually time independent.

@@

We take the bulk metric $g_{\mu\nu}$ to be flat. Let $\{\tau,\sigma\} := \{x^0, x^1\}$ be the spacetime co-ordinates on the worldsheet. Derivatives of the $X$'s in terms of these co-ordinates can be written as:
\begin{equation}
	\dot X := \left\{\frac{\partial X^\mu}{\partial \tau}\right\}; \quad X' := \left\{\frac{\partial X^\mu}{\partial \sigma}\right\}
\end{equation}
 we obtain the following form for the determinant of the worldsheet metric (\eqref{eqn:patch-area-v1}) in terms of the embedding fields:
\begin{align}\label{eqn:area-embedding}
	\det h & = \begin{vmatrix}
		h_{00} && h_{01} \\
		h_{10} && h_{11} 
	\end{vmatrix} = \begin{vmatrix}
		\dot X^2 && \dot X \cdot X' \\
		\dot X \cdot X' && X'^2
	\end{vmatrix} \nonumber \\
	& = \dot X^2 X'^2 - (\dot X \cdot X')^2
\end{align}

@@

We can proceed to obtain the area of a 2-surface from the bulk vielbein as follows. Starting with the bulk vielbein $e_\mu^I$, we perform a local Lorentz transformation such that two ``legs'' of the vielbein are tangent to the worldsheet surface. We then project down the vielbein $e_\mu^I$ to a dyad $e_a^i$ living in the tangent space of the worldsheet, in terms of which the worldsheet metric can be written as:
\begin{equation}
	h_{ab} = e_a^i e_b^j \eta_{ij},
\end{equation}
where $\eta_{ij}$ is the flat $(1+1)$ Minkowski metric. The determinant then becomes:
\begin{align}\label{eqn:area-vielbein}
	\det h & = \begin{vmatrix}
		h_{00} && h_{01} \\
		h_{10} && h_{11} 
	\end{vmatrix} = \begin{vmatrix}
		e_x{}^i e_x{}^j & e_x{}^i e_y{}^j \\
		e_y{}^i e_x{}^j & e_y{}^i e_y{}^j
	\end{vmatrix} \eta_{ij} \nonumber \\
	& = \vec e_x{}^2 \vec e_y{}^2 - (\vec e_x \cdot \vec e_y)^2,
\end{align}
where $\vec e_x \cdot \vec e_y = e_x{}^i e_y{}^j \eta_{ij}$. It is clear by comparing the two expressions \eqref{eqn:area-embedding} and \eqref{eqn:area-vielbein} that we have the following correspondence between the vielbein the $X$ derivatives:
\begin{equation}
	\vec e_\tau \equiv \frac{\partial \vec X}{\partial \tau}; \quad \vec e_\sigma \equiv \frac{\partial \vec X}{\partial \sigma}.
\end{equation}
In other words we can identify the vielbein with the derivatives of the embedding fields. Of course, this is not altogether surprising. If we consider any manifold on which we have some set of fields $\{X^I(x^\mu)\}$, where $I$ is valued in the Lie algebra ${\mc g}$ of some group $\mc G$, as a function of some  ``fiducial'' co-ordinates $\{x^\mu\}$, then we can always define a set of ${\mc g}$ valued vector fields: 
$$
	e^I_\mu := \frac{\partial X^I}{\partial x^\mu}.
$$
In the usual formulation of the string action the embedding fields are \emph{not} Lie-algebra valued but are co-ordinate fields on what is generally taken to be a flat background. In order for this correspondence between vielbein and $X$ derivatives to work we have to view the $X$ fields as taking values in a Lie algebra. But these same $X$ fields \emph{also} serve as local co-ordinates for the bulk. In order for the co-ordinates to change from point to point in the bulk geometry, we must therefore have that the connection $\mc A$ in \eqref{eqn:covariant-deriv-v2} can never vanish completely! If $\mc A$ vanishes then that will imply that the fields $X$ don't change from one point to the next. Thus the amplitude $|\mc A|$ can be arbitrarily small but it can never be zero.

>>>>

In the same way we considered the quantum geometry inverse metric correction to the Nambu-Goto action, it is also natural to consider such correction to the Polyakov string. We start with the action(\ref{eqn:polyakov-gravity}) and proceed with the inverse area correction(\ref{eqn:inverse-metric-correction}), which yields our corrected Polyakov action
\begin{multline} \label{eqn:inverse-metric-Polyakov}
	S_{I A P} = - \frac{T}{2} \int \diff ^2 x \left( \sqrt{- g} - \frac{k \Delta}{2 \sqrt{- g}} \right) \times \\ \times g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu} .
\end{multline}
The reason for considering this form and not one in which the inverse metric $g^{a b}$ also gets a correction term is that this form of a Polyakov inverse area-corrected action reduces nicely to the Nambu-Goto corrected action(\ref{eqn:modified-NG-action}) when the auxiliary metric $g_{a b}$ coincides with the induced metric $h_{a b} = \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu}$. The equations of motion for the auxiliary metric are completely unchanged by this correction, so they are
\begin{equation} \label{eqn:vary-wrt-g}
	\frac{\delta S_{I A P}}{\delta g^{c d}} = 0
\end{equation}
from which we obtain the stress-energy tensor for the worldsheet:
\begin{multline} \label{eqn:inverse-area-em-tensor}
	T_{c d} := \partial _{c} X^{\mu} \partial _{d} X^{\nu} G_{\mu \nu} - \\ - \frac{1}{2} g_{c d} g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu} = 0 ,
\end{multline}
from which we can extract the metric $g_{c d}$ as
\begin{equation} \label{eqn:ws-metric}
	g_{c d} = 2 f \partial _{c} X^{\mu} \partial _{d} X^{\nu} G_{\mu \nu} ,
\end{equation}
where the function $f$ is given by
\begin{equation} \label{eqn:inverse-area-f}
	\frac{1}{f} = g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu} .
\end{equation}

As for variation w.r.t the embbeding fields $X^{\lambda}$, we have
\begin{equation} \label{eqn:vary-wrt-X}
	\frac{\delta S_{I A P}}{\delta X^{\lambda}} = 0
\end{equation}

\begin{multline} \label{eqn:X-eom-inverse-area}
	F^{-} \sqrt{- g} g^{a b} \partial _{a} \partial _{b} X^{\nu} G_{\lambda \nu} + \partial _{a} \left( F^{-} \sqrt{- g} g^{a b} \right) \partial _{b} X^{\nu} G_{\lambda \nu} = \\ = \frac{1}{2} F^{-} \sqrt{- g} g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} \partial _{\lambda} G_{\mu \nu} ,
\end{multline}
where $F^{-} = 1 - k \Delta / 2 (- g)$. In conformal gauge $g^{a b} = (\phi (x))^{-1} \eta ^{a b}$ and assuming flat background $G_{\mu \nu} = \eta _{\mu \nu}$, eq(\ref{eqn:X-eom-inverse-area}) simplify to
\begin{equation} \label{eqn:X-eom-simplified}
	\eta ^{a b} \partial _{a} \partial _{b} X^{\mu} + \eta ^{a b} \partial _{a} \ln (F^{-}) \partial _{b} X^{\mu} = 0 .
\end{equation}

@@

Equation(\ref{eqn:X-eom-simplified}) seems a bit odd, but we can polish it by choosing a plane-wave ansatz for the embedding fields $X^{\mu} = X^{\mu} _{0} \e ^{-i (E \tau - p \sigma)}$ and using it to fix a form for the conformal factor $\phi (x)$ appearing in $F^{-}$. By using the ansatz in eq(\ref{eqn:X-eom-simplified}), it becomes
\begin{equation}
	(E^{2} - p^{2}) X^{\mu} + i (E \partial _{\tau} \ln (F^{-}) - p \partial _{\sigma} \ln (F^{-})) X^{\mu} = 0 ,
\end{equation}
from which we can use the freedom in the conformal factor to fix the second term as $- m^2 X^{\mu}$, with $- m^2 = p^2 - E^2$ on-shell, and this in turn implies that the derivatives of $\ln (F^{-})$ are given by
\begin{equation} \label{eqn:derivatives-F-minus}
	\partial _{\tau} \ln (F^{-}) = i E , \ \partial _{\sigma} \ln (F^{-}) = i p ,
\end{equation}
thus we have
\begin{equation} \label{eqn:ln-F}
	\ln (F^{-}) = i (E \tau + p \sigma) + a , \ a \in \mathbb{C}
\end{equation}

\begin{equation} \label{eqn:F-minus-exp}
	F^{-} = A \e ^{i (E \tau + p \sigma + \varphi)} , \ A > 0
\end{equation}
\begin{equation} \label{eqn:phi}
	\phi (\tau , \sigma) = \sqrt{\frac{k \Delta}{2 \left( 1 - A \e ^{i (E \tau + p \sigma + \varphi)} \right)}} .
\end{equation}
At first glance it seems problematic that we have a complex conformal factor $\phi (x)$, but the complex structure comes from our ansatz for the embedding fields, and when we choose either the real or imaginary part of the latter it will also pick the respective component of $\phi$. With this choice, we thus turned eq(\ref{eqn:X-eom-simplified}) into a Klein-Gordon equation
\begin{equation} \label{eqn:X-Klein-Gordon-eqn}
	\left( \partial _{\sigma} ^2 - \partial _{\tau} ^2 \right) X^{\mu} - \mu ^2 X^{\mu} = 0 .
\end{equation}

@@

At first glance, our recent analysis of the Polyakov string with inverse metric correction seems to not be related to what we found in the analysis of the Nambu-Goto string. However, as pointed out in \cite{Grimshaw2010Homogenization}, wave equations with a variable speed (the result of the Nambu-Goto analysis) can be turned into Klein-Gordon equations (what we just found). While they're work deals only with space-varying wave speed, it is straight forward to generalise the procedure to also include time dependence.

@@

We take the bulk metric $g_{\mu\nu}$ to be flat. Let $\{\tau,\sigma\} := \{x^0, x^1\}$ be the spacetime co-ordinates on the worldsheet. Derivatives of the $X$'s in terms of these co-ordinates can be written as:
\begin{equation}
	\dot X := \left\{\frac{\partial X^\mu}{\partial \tau}\right\}; \quad X' := \left\{\frac{\partial X^\mu}{\partial \sigma}\right\}
\end{equation}
we obtain the following form for the determinant of the worldsheet metric (\eqref{eqn:patch-area-v1}) in terms of the embedding fields:
\begin{align}\label{eqn:area-embedding}
	\det h & = \begin{vmatrix}
		           h_{00} &  & h_{01} \\
		           h_{10} &  & h_{11}
	           \end{vmatrix} = \begin{vmatrix}
		                           \dot X^2        &  & \dot X \cdot X' \\
		                           \dot X \cdot X' &  & X'^2
	                           \end{vmatrix} \nonumber \\
	       & = \dot X^2 X'^2 - (\dot X \cdot X')^2
\end{align}

@@

We can proceed to obtain the area of a 2-surface from the bulk vielbein as follows. Starting with the bulk vielbein $e_\mu^I$, we perform a local Lorentz transformation such that two ``legs'' of the vielbein are tangent to the worldsheet surface. We then project down the vielbein $e_\mu^I$ to a dyad $e_a^i$ living in the tangent space of the worldsheet, in terms of which the worldsheet metric can be written as:
\begin{equation}
	h_{ab} = e_a^i e_b^j \eta_{ij},
\end{equation}
where $\eta_{ij}$ is the flat $(1+1)$ Minkowski metric. The determinant then becomes:
\begin{align}\label{eqn:area-vielbein}
	\det h & = \begin{vmatrix}
		           h_{00} &  & h_{01} \\
		           h_{10} &  & h_{11}
	           \end{vmatrix} = \begin{vmatrix}
		                           e_x{}^i e_x{}^j & e_x{}^i e_y{}^j \\
		                           e_y{}^i e_x{}^j & e_y{}^i e_y{}^j
	                           \end{vmatrix} \eta_{ij} \nonumber        \\
	       & = \vec e_x{}^2 \vec e_y{}^2 - (\vec e_x \cdot \vec e_y)^2,
\end{align}
where $\vec e_x \cdot \vec e_y = e_x{}^i e_y{}^j \eta_{ij}$. It is clear by comparing the two expressions \eqref{eqn:area-embedding} and \eqref{eqn:area-vielbein} that we have the following correspondence between the vielbein the $X$ derivatives:
\begin{equation}
	\vec e_\tau \equiv \frac{\partial \vec X}{\partial \tau}; \quad \vec e_\sigma \equiv \frac{\partial \vec X}{\partial \sigma}.
\end{equation}
In other words we can identify the vielbein with the derivatives of the embedding fields. Of course, this is not altogether surprising. If we consider any manifold on which we have some set of fields $\{X^I(x^\mu)\}$, where $I$ is valued in the Lie algebra ${\mc g}$ of some group $\mc G$, as a function of some  ``fiducial'' co-ordinates $\{x^\mu\}$, then we can always define a set of ${\mc g}$ valued vector fields:
$$
	e^I_\mu := \frac{\partial X^I}{\partial x^\mu}.
$$
In the usual formulation of the string action the embedding fields are \emph{not} Lie-algebra valued but are co-ordinate fields on what is generally taken to be a flat background. In order for this correspondence between vielbein and $X$ derivatives to work we have to view the $X$ fields as taking values in a Lie algebra. But these same $X$ fields \emph{also} serve as local co-ordinates for the bulk. In order for the co-ordinates to change from point to point in the bulk geometry, we must therefore have that the connection $\mc A$ in \eqref{eqn:covariant-deriv-v2} can never vanish completely! If $\mc A$ vanishes then that will imply that the fields $X$ don't change from one point to the next. Thus the amplitude $|\mc A|$ can be arbitrarily small but it can never be zero.
